\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}

\begin{document}

\section{Introduction}

Recall questions of origin of adaptations

Reference on parallel adapation, standing variation, etc.:
 coop,
 pennings and hermisson,

Summarize previous paper

Question assumptions of no standing variation and homogenous selection

Stick to constant-speed waves.

\section{Standing variation}

\subsection{Model}

selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous

For now we ignore geographic hetergeneity, and assume that the species range
is a homogenous, two-dimensional region.  (results for a one-dimensional region also appear)
There are two selective types -- the neutral type, and the mutated type.
We assume that different mutations are distinguishable --
either as selectively equivalent mutations, or by linked variation.
At first, the mutated type is at a selective disadvantage, for long enough to be at selection-mutation equilibrium,
but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
After fixation, alleles are either descended
from families of mutants present as standing variation when the selective regime changed,
or from new mutants arising since that time.

We will suppose that before time $t=0$,
the mutant type has fitness $1-s_d$ relative to the neutral type,
and that after time $t=0$,
the mutant type has fitness $1+s_b$,
for positive numbers $s_d$ and $s_b$.
We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
(If the mutation is recessive, XXX more complicated things happen.)

For the other parameters,
suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
and that the XXX dispersal variance is $\sigma^2$.
The species occupies an area with mean density
$\rho$ chromosomes per unit area.
(XXX or do diploid? below is haploid.)

We make use of the commonly used approximation
that neglects spatial structure and competition between relatives,
treating the offspring of a new mutant that appears in an area not already occupied by the mutated type
as a (nonspatial) branching process.
After $t=0$, the offspring of a single mutant form (approximately) a branching process with growth rate $s_b$,
so each new mutant establishes locally with probability $p_s \approx 2s_b$.
Therefore, the total rate at which new mutants appear and establish locally after $t=0$ in areas not already occupied by the mutant type
is approximately $\lambda = 2 \mu \rho s_b$.
Before $t=0$, the mutation is deleterious, so the genetic descendents of each new mutation are doomed to extinction,
but may persist for some time.
Consider a mutation that arose at time $-T<0$, and let $Z_s$ be the number of its descendants at time $s-T$.
At time $t=0$, there are $Z_T$ individuals present with the mutation,
and each has probability $p_s \approx 2s_b$ of establishing, approximately independently.
Therefore, the probability that some descendants of this mutation establish and fix locally is
$1-(1-p_s)^Z_T$, XXX continue here

If $Z_t$ is a branching process with $Z_0 = 1$ and $\E[Z_t] = \exp(-s_d t)$, then



For $t \ge 0$, new mutations appear at rate approximately $\lambda = 4 \mu \rho s_b$.

rescale to two parameters
-- a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation

Branching-process-sized point mass approx to local groups present at $t=0$
-- depends on local population density large enough relative to migration (to be "branching")
-- and migration not too large (since we approximate as a point mass).
Find mean density of clusters of successful standing mutants. 
Is greater-than-one size of clusters likely to make a difference? 
 
recall PPP computations; draw cones of opportunity for new mutations 
 
\subsection{Results} 
 
proportion of haplotypes (or of space?) that are from standing variation 
 
mean (and variance?) of distance between haplotypes 
 
influence of migration on things -- plot these versus migration 
 
 
\section{Patchy selection} 
 
\subsection{Model} 
 
patches where selection is $1+s_b$ separated by $1-s_d$ -- separation large relative to migration.   
(or relative to $s_d/m$?) 
 
Describe simple discrete model. 
 
Approximation in continuous model by branching Brownian motion with killing to find migration rate between patches -- 
good approximation neglecting interaction (if local population size is large). 
Look at Barton. 
 
\subsection{Results} 
 
Pull out results from other paper on discrete model. 
 
when are patches effectively isolated? 
 
\section{Applications} 
 
Look at some human parameters? 
 
\section{Discussion} 
 
if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
recall others' work. 
 
patchy selection helps this happen; 
effective migration rate looks like what; 
strength of (positive?) selection suprisingly does not affect results (?) 
 
standing variation will be important under these circumstances; 
migration rate (and so geographic structure) will not affect numbers of types if this is true; 
however migration always affcts spatial patterns. 
 
discussion of applications. 
 
\end{document} 
